---
title: "Lab 4 Homework"
author: "Dawson Diaz"
date: "February 7th, 2020"
output:
  html_document: 
    keep_md: yes
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Problems

The data used for this assignment are from the [United Nations Food and Agriculture Organization](http://www.fao.org/about/en/) on world fisheries. 
<br>
<br>
***1. Load the data `FAO_1950to2012_111914.csv` as a new object titled `fisheries`.***


```{r message=FALSE, warning=FALSE}
fisheries <- readr::read_csv("data/FAO_1950to2012_111914.csv")
```

<br>

***2. What are the names of the columns? Do you see any potential problems with the column names?***

```{r}
colnames(fisheries)
```
There are a couple of potential problems with the column names, (1) some names have spaces, (2) there are data contained in the columns-- there are many individual columns that are levels of the variable "year" that should just be one column.

<br>


***3. What is the structure of the data? Show the classes of each variable.***

This data is structured in wide format. While this is perhaps easiest to enter in data, it creates problems for analysis. The classes of each variable are shown usinf the `lapply()` function.

```{r}
lapply(fisheries, class)
```


<br> 
***4. Think about the classes. Will any present problems? Make any changes you think are necessary below.***

The values of each year, which are numbers are all characterized as `character` when they should be `numeric`. This can be more efficicently fixed later when the data is tidy.
Meanwhile other columns are `characters` when they would be better treated as a `factor`


```{r}
fisheries$Country <- as.factor(fisheries$Country)
fisheries$`ISSCAAP taxonomic group` <- as.factor(fisheries$`ISSCAAP taxonomic group`)
fisheries$`ASFIS species#` <- as.factor(fisheries$`ASFIS species#`)
fisheries$`FAO major fishing area` <- as.factor(fisheries$`FAO major fishing area`)
fisheries$`Common name` <- as.factor(fisheries$`Common name`)
fisheries$`ASFIS species name` <- as.factor(fisheries$`ASFIS species name`)
```


<br>

***5. How many countries are represented in the data? Provide a count.***

```{r}
# Get the number of levels of the column variable Country
nlevels(fisheries$Country)
```


<br>

***6. What are the names of the countries?***

```{r}
levels(fisheries$Country)
```

<br>
***7. Use `rename()` to rename the columns and make them easier to use.***

```{r}
fisheries <- fisheries %>% 
  
  dplyr::rename(
    country = Country,
    commname = `Common name`,
    ASFIS_sciname = `ASFIS species name`,
    ASFIS_spcode = `ASFIS species#`,
    ISSCAAP_spgroup = `ISSCAAP group#`,
    ISSCAAP_spgroupname = `ISSCAAP taxonomic group`,
    FAO_area = `FAO major fishing area`,
    unit = Measure
  )
```

<br>

***8. Are these data tidy? Why or why not, and, how do you know? Use the appropriate pivot function to tidy the data. Remove the NA's. Put the tidy data frame into a new object `fisheries_tidy`.***  

These data are not tidy. 

```{r}
fisheries_tidy <- fisheries %>% 
  pivot_longer(
    cols= "1950":"2012",
    names_to = "year",
    values_to = "catch_tonnes",
    values_drop_na = TRUE
  )
```


<br>
***9. Refocus the data only to include country, ISSCAAP_spgroupname, ASFIS_spcode, ASFIS_sciname, year, and catch.***

```{r}
fisheries_tidy2 <- fisheries_tidy %>% 
  select(country, ISSCAAP_spgroupname, ASFIS_spcode, ASFIS_sciname, year, catch_tonnes)
```

```{r}
fisheries_tidy2
```


<br>
***10. Re-check the classes of `fisheries_tidy2`. Notice that "catch" is shown as a character! This is a problem because we will want to treat it as a numeric. How will you deal with this?***


```{r warning=FALSE}
fisheries_tidy2$catch_tonnes <- as.numeric(fisheries_tidy2$catch_tonnes)
```
```{r}
lapply(fisheries_tidy2, class)
```

<br>

***11. Based on the ASFIS_spcode, how many distinct taxa were caught as part of these data?***

```{r}
n_distinct(fisheries_tidy2$ASFIS_spcode)
```


<br>

***12. Which country had the largest overall catch in the year 2000?***


```{r}
global_total_catch_2000 <- fisheries_tidy2 %>% 
  group_by(country) %>%
  filter(year==2000) %>% 
  summarize (
    total_catch = sum(catch_tonnes, na.rm=TRUE) 
  ) %>% 
  arrange(desc(total_catch))
  
```

```{r}
head(global_total_catch_2000)
```
In 2000, Peru was the country with the largest total catch (in tonnes).

<br>


***13. Which country caught the most sardines (_Sardina_) between the years 1990-2000?***

```{r}
fisheries_tidy2$year <- as.numeric(fisheries_tidy2$year)
fisheries_tidy2$ASFIS_sciname <- as.character(fisheries_tidy2$ASFIS_sciname)
```

```{r}
sardine_total <- fisheries_tidy2 %>% 
  select(year, ASFIS_sciname, ISSCAAP_spgroupname, country, catch_tonnes) %>%
  filter(year >=1990, year<=2000, ASFIS_sciname=="Sardina pilchardus") %>% 
  group_by(country) %>% 
  summarize(sardina_catch_tonnes = sum(catch_tonnes, na.rm=TRUE)) %>% 
  arrange(desc(sardina_catch_tonnes))
```

```{r}
head(sardine_total)
```
Between 1990 and 2000, Morocco caught the most Sardina pilchardus with 4785190 tonnes caught.



<br>
***14. Which five countries caught the most cephalopods between 2008-2012?***

```{r}
 cephalopod_total <- fisheries_tidy2 %>% 
  select(country, year, ISSCAAP_spgroupname, catch_tonnes) %>% 
  filter(ISSCAAP_spgroupname == "Squids, cuttlefishes, octopuses") %>% 
  filter(year >=2008, year <=2012) %>% 
  group_by(country) %>% 
  summarize(cephalopod_catch_tonnes = sum(catch_tonnes, na.rm=TRUE)) %>% 
  arrange(desc(cephalopod_catch_tonnes))
```

```{r}
head(cephalopod_total)
```

Between 2008 and 2012, China caught the most cepalopods with 4785139 tonnes caught.


<br>
***15. Given the top five countries from question 12 above, which species was the highest catch total? The lowest?***

The top five countries with the highest global catch in 2000 are: Peru, Japan, The United States, Chile, and Indonesia. 

```{r}
top5 <- c("Peru", "Japan","United States of America", "Chile", "Indonesia")
```


```{r}
top5_catch <- fisheries_tidy2 %>% 
  select(country, year, ASFIS_sciname, catch_tonnes) %>% 
  filter(year==2000) %>% 
  filter(country %in% top5) %>% 
  group_by(ASFIS_sciname) %>% 
  summarize(
    catch_tonnes = max(catch_tonnes, na.rm=TRUE)
  ) %>% 
  arrange(desc(catch_tonnes))
```

Among the top 5 countries with the highest catch in 2000, `Engraulis ringens` was the most caught fish.
```{r}
top_n(top5_catch, 5, catch_tonnes)
```


```{r}
top_n(top5_catch,-19, catch_tonnes)
```

Among the top 5 countries with the highest catch in 2000, `Tresus spp`, `Tawera gayi`, `Pleuronichthys decurrens`, `Lobotes surinamensis`, `Caulolatilus microps`, `Carcharodon carcharias` `Carcharias taurus` and `Brotula barbata` were all tied for the least caught.


<br>
***16. In some cases, the taxonomy of the fish being caught is unclear. Make a new data frame called `coastal_fish` that shows the taxonomic composition of "Miscellaneous coastal fishes" within the ISSCAAP_spgroupname column.***


```{r}
coastal_fish <- fisheries_tidy2 %>% 
  select(ISSCAAP_spgroupname, ASFIS_sciname) %>% 
  filter(ISSCAAP_spgroupname == "Miscellaneous coastal fishes") %>% 
  pivot_longer(cols = ASFIS_sciname,
               names_to = "species",
               values_to= "taxa"
  )
```

```{r warning=TRUE}
coastal_fish$taxa <- as.factor(coastal_fish$taxa)
```

```{r}
nlevels(coastal_fish$taxa)
```


Miscellaneous coastal fishes  are comprised of the following `412` species:
```{r}
levels(coastal_fish$taxa)
```


<br>
***17. Use the data to do at least one exploratory analysis of your choice. What can you learn?***

Following the Fukishima nuclear diseaser in Japan did fishing levels immediately drop? -- Since data only goes until 2012, I will compare 2012 levels to 2011 and 2010 levels.

```{r}
fisheries_tidy2 %>% 
  select(country, year, ASFIS_sciname, catch_tonnes) %>% 
  filter(country=="Japan") %>% 
  filter(year <=2012, year>=2010) %>% 
  group_by(year) %>% 
  summarize(
    total_catch_japan = sum(catch_tonnes, na.rm=TRUE)
  )
```

Wow! The fishing industry seems to have been really affected by the nucelear diseaser.



<br><br><br><br>