---
title: "Lab 5 Homework"
author: "Dawson Diaz"
date: "February 14th, 2020"
output:
  html_document: 
    keep_md: yes
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Libraries
```{r message=FALSE}
library(tidyverse)
library(naniar)
library(skimr)
```

### Data

### amniota data
Myhrvold N, Baldridge E, Chan B, Sivam D, Freeman DL, Ernest SKM (2015). “An amniote life-history
database to perform comparative analyses with birds, mammals, and reptiles.” _Ecology_, *96*, 3109.
doi: 10.1890/15-0846.1 (URL: https://doi.org/10.1890/15-0846.1).
```{r message=FALSE, warning=FALSE}
amniota <- 
  readr::read_csv("data/amniota.csv")
```

### amphibio data
Oliveira BF, São-Pedro VA, Santos-Barrera G, Penone C, Costa GC (2017). “AmphiBIO, a global database
for amphibian ecological traits.” _Scientific Data_, *4*, 170123. doi: 10.1038/sdata.2017.123 (URL:
https://doi.org/10.1038/sdata.2017.123).
```{r message=FALSE, warning=FALSE}
amphibio <- 
  readr::read_csv("data/amphibio.csv")
```

### Questions  
**1. First, do some exploratory analysis of both data sets. What is the structure, column names, and dimensions?**
```{r}
glimpse(amniota)
```
<br>
```{r}
glimpse(amphibio)
```
<br>
**2. How many total NAs are in each data set? Do these values make sense? Are NAs represented by any other values?**

In the `amniota` dataset, NA values are represented as `-999` so they are not picked up by the `is.na()` function.

```{r}
amniota %>% 
  summarize(total_na = sum(is.na(amniota)))
```

In the `amphibio` dataset, NA values are represented as `NA`.
<br>
```{r}
amphibio %>% 
  summarize(total_na = sum(is.na(amphibio)))
```


<br>

**3. Make any necessary replacements in the data such that all NAs appear as "NA".**

We only need to fix the `amniota` dataset using the following command:
```{r}
amniota <- amniota %>% 
  na_if("-999")
```



<br>
**4. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for both data sets.**

```{r}
naniar::miss_var_summary(amphibio)
```
<br>
```{r}
naniar::miss_var_summary(amniota)
```


<br>

**5. For the `amniota` data, calculate the number of NAs in the `egg_mass_g` column sorted by taxonomic class; i.e. how many NAs are present in the `egg_mass_g` column in birds, mammals, and reptiles? Does this results make sense biologically? How do these results affect your interpretation of NAs?**

```{r message=FALSE, warning=FALSE}
amniota %>% 
  select(class, egg_mass_g) %>% 
  group_by(class)  %>% 
  naniar::miss_var_summary()
```

It makes sense that Mammalia would have most/all its `egg_mass_g` data missing since most mammals do not lay eggs. However, I would expect that Aves would have more of its data present.

<br>

**6. Which taxonomic classes are represented in the data? Summarize this in a chart then make a barplot that shows the proportion of observations in each taxonomic class.**

Birds, Mammals and Reptiles are represented in the `amniota` dataset.

```{r}
amniota %>% 
  group_by(class) %>% 
  ggplot(mapping = aes(x=class))+geom_bar(stat="count")
```



<br>

**7. Let's explore the taxonomic composition of the data a bit more. How many genera are represented by taxonomic class? Present this as a data table and bar plot.**

```{r}
amniota_genera_by_class <- amniota %>% 
  group_by(class) %>%
  summarise(
   genera = n_distinct(genus))
```

```{r}
amniota_genera_by_class
```

<br>

```{r}
ggplot(amniota_genera_by_class,  mapping=aes(x=class, y=genera))+geom_bar(stat="identity")
```



<br>

**8. Lastly, how many species are represented in each taxonomic class? Present this as a data table and bar plot.**

```{r}
amniota_species_by_class <- amniota %>% 
  group_by(class) %>% 
  summarise(species = n_distinct(species))
```

```{r}
amniota_species_by_class
```


```{r}
ggplot(amniota_species_by_class,  mapping=aes(x=class, y=species))+geom_bar(stat="identity")
```


**9. The `amphibio` data includes information on ecology. Each species is classified as fossorial, terrestrial, aquatic, or arboreal. How many species are in each of these categories? Make a bar plot to visualize these results. Which category is most/ least represented in the data?**

```{r}
amphibio_by_eco <- amphibio %>% 
  summarize(
    fossorial = sum(Fos, na.rm=TRUE),
    terrestrial= sum(Ter, na.rm=TRUE),
    aquatic= sum(Aqu,na.rm=TRUE),
    arboreal= sum(Arb,na.rm=TRUE)) %>% 
  pivot_longer(fossorial:arboreal,
    names_to = "ecology",
    values_to = "count"
  )
```

```{r}
amphibio_by_eco
```


```{r}
ggplot(amphibio_by_eco, mapping = aes(x=ecology, y=count))+geom_bar(stat="identity")
```

Terrestrial amphibians are the most represented whilist fossorial are the least represented.


<br>
**10. For the amphibio data, we are interested to know if body size is correlated with litter size? Make a plot that shows this relationship. You should notice that there are outliers. Can you think of a way to identify the outliers for both body size and litter size? What happens when you then re-plot the data?**

```{r warning=FALSE}
amphibio %>%
  select(Body_size_mm, Litter_size_min_n, Litter_size_max_n) %>%
  rowwise() %>% mutate(Litter_size_avg_n = mean(c(Litter_size_min_n, Litter_size_max_n),  na.rm=TRUE)) %>% 
  ggplot(amphibio_plot, mapping=aes(x=Body_size_mm, y=Litter_size_avg_n))+geom_jitter()
```

<br>
We can filter out  all values of `litter_size_avg_n` above 1000 to see if any relationship exists. It would  appear as if no clear relationship exists.


```{r warning=FALSE}
amphibio %>% 
  select(Litter_size_min_n, Body_size_mm) %>% 
  filter(Litter_size_min_n < 1000) %>% 
  ggplot(mapping=aes(x=Body_size_mm, y=Litter_size_min_n))+geom_jitter()+geom_smooth(method=lm, se=T)
```
<br><br><br>






