---
title: "Lab 3 Homework"
author: "Dawson Diaz"
date: "Winter 2020"
output:
  html_document: 
    keep_md: yes
    theme: spacelab
---
<br>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Load the tidyverse**

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

<br>

**1. Load the data into a new object called `homerange`.**  

```{r message=FALSE, warning=FALSE}
homerange <- readr::read_csv("data/Tamburelloetal_HomeRangeDatabase.csv")
```

<br>

**2. Use `spec()` to see the full details of the columns.**

```{r}
spec(homerange)
```

<br>

**3. Explore the data. Show the dimensions, column names, classes for each variable, and a statistical summary. Keep these as separate code chunks. **

Dimensions
```{r}
dim(homerange)
```
<br>
Column Names
```{r}
colnames(homerange)
```
<br>
Classes

the  `lapply()` function applies the `class()` function to the dataframe columns of `homerange`.
```{r}
lapply(homerange, class)
```
<br>
Stastical Summaries
```{r}
summary(homerange)
```

<br>

**4. Are there NA's in your data? Show the code that you would use to verify this, please.**

Yes, there are NA values in the data. This is verified by running the `anyNA` function.
```{r}
anyNA(homerange)
```

<br>
**5. Change the class of the variables `taxon` and `order` to factors and display their levels.**

```{r}
homerange$taxon <- as.factor(homerange$taxon)
levels(homerange$taxon)
```

<br>

```{r}
homerange$order <- as.factor(homerange$order)
levels(homerange$order)
```

<br>


**6. Make a new dataframe `deer` that is limited to the mean mass, log10 mass, family, genus, and species of deer in the database. The family for deer is cervidae. Arrange the data in descending order by log10 mass. Which is the largest deer?** 

```{r}
deer <-homerange %>%
  select(family, genus, species, mean.mass.g, log10.mass) %>% 
  filter(family == "cervidae") %>% 
  arrange(desc(log10.mass))
```

```{r}
deer
```

<br>

To find the largest deer...
```{r}
largest_deer <- max(deer$mean.mass.g, na.rm=TRUE)
filter(deer, mean.mass.g == largest_deer)
```

<br>

**7. As measured by the data, which snake species has the smallest homerange? Show all of your work, please. Look this species up online and tell me about it!**  

```{r}
snakes <- homerange %>% 
    select(taxon, family, genus, species, mean.hra.m2) %>% 
    filter(taxon == "snakes") %>% 
    arrange(mean.hra.m2)
```
<br>
```{r}
min_hra.m2_snake <- min(snakes$mean.hra.m2)
filter(snakes, mean.hra.m2 == min_hra.m2_snake)
```
The snake species with the smallest homerange is *Bitis schenideri*. Found in the coastal regions of Namibia and South Africa, it has an average homerange of 200 square meters. It has an average length of 10-25cm, which might make it the world's smallest viperid. It was named after Dr. Oskar Schneider.
<br><br>
**8. You suspect that homerange and mass are correlated in birds. We need a ratio that facilitates exploration of this prediction. First, build a new dataframe called `hra_ratio` that is limited to genus, species, mean.mass.g, log10.mass, log10.hra. Arrange it in ascending order by mean mass in grams.**

```{r}
hra_ratio <- homerange %>% 
  filter(taxon=="birds") %>% 
  select(genus, species, mean.mass.g, log10.mass, log10.hra) %>% 
  arrange(mean.mass.g)
```

```{r}
hra_ratio
```


<br>
**9. Replace the existing `hra_ratio` dataframe with a new dataframe that adds a column calculating the ratio of log 10 hra to log 10 mass. Call it `hra.mass.ratio`. Arrange it in descending order by mean mass in grams.** 


```{r}
hra_ratio <- hra_ratio %>%
  mutate(hra.mass.ratio = log10.hra/log10.mass) %>% 
  arrange(desc(mean.mass.g))
```


```{r}
hra_ratio
```


<br>
**10. What is the lowest mass for birds with a `hra.mass.ratio` greater than or equal to 4.0?**

```{r}
bird_hra_mass_ratio <- hra_ratio %>% 
  filter(hra.mass.ratio >= 4.0) %>% 
  arrange(mean.mass.g)
```

```{r}
bird_hra_mass_ratio
```


The lowest mass is:
```{r}
min(bird_hra_mass_ratio$mean.mass.g)
```


<br>
**11. Do a search online; what is the common name of the bird from #10 above? Place a link in your markdown file that takes us to a webpage with information on its biology.**  


The bird with the lowest mass with a hra:mass ratio of >= 4.0 is *Regulus regulus* Its common name is the Goldcrest.<br> 
To find out more click [here](https://avibase.bsc-eoc.org/species.jsp?avibaseid=A7CC62D6BD333F25)


<br>
**12. What is the `hra.mass.ratio` for an ostrich? Show your work, please.**  


```{r}
hra_ratio %>% 
  select(genus, species, hra.mass.ratio) %>% 
  filter(genus=="struthio")
```



<br><br><br> 